---
title: "stringr Pattern Matching tasks"
author: "Andrea Chandler"
date: "2/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("stringr")
library(stringr)
```


We will be going over pattern matching function tasks, which allow stringr to detect, locate, extract, match, replace and split strings. 


Lets get started!

First, lets set up some example strings. In this one, you have a collection of phone numbers and random words.
```{r}
strings <- c("apple", "219 733 8965", "grape", "329-293-8753", "Work: 579-499-7527; Home 543.355.3679")

phone <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"
#this variable is designed to match phone numbers 
```
First, we'll go through detect and subset

str_detect() detects the presence or absence of a pattern and returns a logical vector 

str_subset() returns the elements of a character vector that match a regular expression 

```{r}
#detect just tells you which strings contain phone numbers and returns it in a logical vector

str_detect(strings, phone)
```

```{r}
#Str_subset actually returns the elements that we specified in the variable phone, in this case phone numbers

str_subset(strings, phone)
```

let's practice detect and subset on a new string set of common names
```{r}
strings2 <- c("joseph", "jt", "michael", "ruchira", "mary", "alex", "jessica", "andrea", "joe", "stella")

names <- c("a")

```

What if you wanted to know the number of 'a's in each name?
```{r}
#str_count() counts the number of matches
str_count(strings2, names)

#try it with our original phone number!

```

What if we wanted to know where the 'a's were in each name?
```{r}
#str_locate() locates the first position of a pattern (the letter a) and returns a numeric matrix with start and end

str_locate(strings2,names)
#this one makes a little more sense with the phone number example

#str_locate_all() locates ALL the positions of the pattern, not just the first

str_locate_all(strings2, names)
```

Extracting the text that matches our defined pattern

```{r}
#extracts the first match
str_extract(strings, phone)

#how do you think we would extract ALL matches, not just the first?
```

str_match() is alittle different. It extracts groups seperated by parenthesis () in our pattern. Since our names example does not have () in the pattern, we have to use our phone example

```{r}
#str_match() extracts groups formed by () from the first match
str_match(strings, phone)

#str_match() extracts groups formed by () from all matches
```

```{r}
#replacing all the phone numbers with XXX-XXX-XXX
str_replace(strings, phone, "XXX-XXX-XXX")

#there's also a replace all function!


#replacing all the 'a's with 'e's 

```

Let's cheat in Wordle!

Here we have a random selection of all the possible wordle words.

```{r}
words <- c("cigar", "rebut", "sissy", "humph", "awake", "blush", "focal","evade", "naval", "serve", "heath", "dwarf", "model", "karma","stink", "grade", "quiet", "bench", "abate", "feign", "major","death", "fresh", "crust", "stool", "colon", "abase", "marry","react", "batty", "pride", "floss", "helix", "croak", "staff", "paper", "unfed", "whelp", "trawl", "outdo", "adobe", "crazy", "sower", "repay", "digit", "crate", "cluck", "spike", "mimic", "pound", "maxim", "linen", "unmet", "flesh", "booby", "forth", "first", "stand", "belly", "ivory", "seedy", "print", "yearn", "drain", "bribe", "stout", "panel", "crass", "flume", "offal", "agree", "error", "swirl", "argue", "bleed", "delta", "flick", "totem", "wooer", "front", "shrub", "parry", "biome", "lapel", "start", "greet", "goner", "golem", "lusty", "loopy", "round", "audit", "lying", "gamma", "labor", "islet", "civic", "forge", "corny", "moult", "basic", "salad", "agate", "spicy", "spray", "essay", "fjord", "spend", "kebab", "guild", "aback", "motor", "alone", "hatch", "hyper", "thumb", "dowry", "ought", "belch", "dutch", "pilot", "tweed", "comet", "jaunt", "enema", "steed", "abyss", "growl", "fling", "dozen", "boozy", "erode", "world", "gouge", "click", "briar", "great", "altar", "pulpy", "blurt", "coast", "duchy", "groin", "fixer", "group", "rogue", "badly", "smart", "pithy", "gaudy", "chill", "heron", "vodka", "finer", "surer", "radio", "rouge", "perch", "retch", "wrote", "clock", "tilde", "store", "prove", "bring", "solve", "cheat", "grime", "exult", "usher", "epoch", "triad", "break", "rhino", "viral", "conic", "masse", "sonic", "vital", "trace", "using")

```




