---
title: "Stringr tutorial Characters + Whitespace Template"
author: "Ruchira Krishnamurthy"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# If you don't have stringr installed
# install.packages("stringr")

# Load stringr
library(stringr)
```

# Getting and Setting Individual Characters 
## str_length() returns the length of a string 
What's your major?
```{r}
str_length()
```


# str_sub() allows you to access individual characters in a string
Name three of your favorite movies that are 1 word
### 3 arguments: str_sub(string, start, end)
```{r}
x <- c()

# Return the 3rd character of each string in the character vector
str_sub()
```

### You can also use str_sub() to modify strings
```{r}
# Replace the 3rd character of each string with "X"
str_sub()<- "X"
x
```

# Add Whitespace
Use str_pad() "pads" a string to a fixed length by adding extra white space on left, right or both sides
### Arguments: str_pad(string, width, side = c("left", "right", "both"))
- Default str_pad() adds space to the left

Name two of your favorite colors.
```{r}
y <- c()

# Pad string to change width to be width of 10 
str_pad()

# Pad the string on both sides
```


# Remove Whitespace 
Use str_trim() to remove leading and trailing white space 
## Arguments: str_trim(string, side = c("both", "left", "right"))
- Default str_trim() removes space from both sides
```{r}
z <- c("  a   ", "b   ",  "   c")

# Trim whitespace on both sides
str_trim()

# Trim whitespace on only left side

```










