---
title: "dplyr-Select, Group By, Summarize"
author: "Ruchira Krishnamurthy"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# dplyr is a package in the tidyverse!
# install.packages("palmerpenguins")
# install.packages("tidyverse")
library(tidyverse)
library(palmerpenguins)
```

# Select
- select() allows us to pick or exclude certain variables in your data

# First let's look at the data set using head()
```{r view dataset}
# View the first 6 rows
head(penguins)
```

# Let's say we want to only view the columns 'year', 'island', and 'species' from the penguin data and in that order.
```{r}
penguins %>%
  select(year, island, species)
```

# View data set using head()
```{r}
# Note: the penguins data set does not change. Using %>% and select() we are choosing what columns or variables we would like to view. 
head(penguins)
```

# View all the columns from species to bill_depth_mm as well as year

```{r}
# : operator
penguins %>%
  select(species:bill_depth_mm, year)
```

# Keep all the columns except the ones from species to bill_depth_mm and year
```{r}
# : operator + !or - operator

# ! operator
penguins %>%
  select(!c(species:bill_depth_mm, year))

# OR - operator
penguins %>%
  select(-c(species:bill_depth_mm, year))
```


# Group_by + Summarize 
Use group_by() and summarize() to prepare a summary table of your data. 
- Group_by() takes an existing table and converts it into a grouped table where operations are performed "by group"
- Summarize() creates a new table containing a statistical summary of data you are interested in

# Example
Generate a summary table containing the mean of penguin body mass (body_mass_g), grouped by penguin species. Column in the summary table should be called mass_mean.
```{r}
# Note that the na.rm = TRUE argument is added to exclude missing values.
penguins %>%
  group_by(species) %>%
  summarize(mass_mean = mean(body_mass_g, na.rm = TRUE))
```

What other statistical operations can we do on this data?
```{r}
penguins %>%
  group_by(species) %>%
  summarize(mass_mean = mean(body_mass_g, na.rm = TRUE),
            mass_median = median(body_mass_g, na.rm = TRUE),
            mass_sd = sd(body_mass_g, na.rm = TRUE))
```

What other variables can we group by? 
```{r}
penguins %>%
  group_by(island, sex) %>%
  summarize(mass_mean = mean(body_mass_g, na.rm = TRUE),
            mass_median = median(body_mass_g, na.rm = TRUE),
            mass_sd = sd(body_mass_g, na.rm = TRUE))
```

What happens if we didn't use the group_by() function?
```{r}
penguins %>%
  summarize(mass_mean = mean(body_mass_g, na.rm = TRUE),
            mass_median = median(body_mass_g, na.rm = TRUE),
            mass_sd = sd(body_mass_g, na.rm = TRUE))
```

Useful links:
https://dplyr.tidyverse.org/reference/summarise.html#useful-functions
https://allisonhorst.shinyapps.io/dplyr-learnr/#section-welcome



