---
title: "Packages and Things We Missed"
author: "Joe Celebrezze"
date: "10/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages

## What are packages?
Packages are groups of functions that typically relate to a certain goal. Anyone can write packages, but not all packages are on CRAN (comprehensive R archive network). As of now, there *are* over 10,000 packages on CRAN. Before we work with functions on specific packages, we must load the packages.

## Installing, updating, removing loading packages that are on CRAN
R and RStudio have established functionality for dealing with packages. To demonstrate this, we will be working with a very useful package called the 'tidyverse' which actually consists of multiple packages including ggplot2 (a data visualization package) and dplyr (a data wrangling/tidying package).

Installing:
```{r}
install.packages('tidyverse')
    # Note: the name of the package must be in quotes; both 'package-name' and "package-name"        would work
```

Updating:
```{r}
update.packages('tidyverse')
```

Removing:
```{r}
remove.packages('tidyverse')
```

Loading:
```{r}
# library('tidyverse')
    # Note: the above line of code will not work because we just removed 'tidyverse' from our        library. In order for this to run, we must install it again

install.packages('tidyverse')
library('tidyverse')
```

Now, since the package has been installed and loaded, we could use the functions on that package!

## Another way to load packages
In almost all cases, there are multiple ways to do the same thing on R. This is true for loading packages. Another way to load packages is to go to the bottom right panel of RStudio and go to packages and select the empty box to the left of the package name to load it. Deselect the checked boxes if you wish to detach a certain package.

## Checking what packages we have
```{r, results='hide'}
installed.packages()
```

## What about packages not on CRAN?
For packages that are not on CRAN, we could still install and load them from the package-makers' GitHub page. This is not entirely common, as there are so many packages available on CRAN; however, for niche things, it is important to know that this is a possibility. See https://support.datacamp.com/hc/en-us/articles/360006286174-Installing-R-Packages-Not-Hosted-on-CRAN for more details.

## Let's explore the dplyr package a bit more.
To explore the dplyr package, we can look at packages (bottom right corner) and click on 'dplyr'. This help page shows all functions included in the dplyr package. Read through some of them, and tell them which function(s) look interesting or useful.

Some of the most useful functions on dplyr:
filter()
group_by()
mutate()
rename()
summarize()
ungroup()

# Reading in Dataframes

## Reading in .csv files
read.csv() is the function to read in a .csv file. Make sure that:

1. The .csv file is located in your *working directory*. The working directory is where R will look for files you want to load and store files you want to save.

  Questions about working directory?
  
2. There are quotes around the name of the file
3. No typos!
    

```{r}
read.csv('sea-urchin.csv')
```


## What about an Excel spreadsheet?
Typically, I would convert an Excel spreadsheet into a .csv file to load it into R. This is the best way that I know to work with data stored on an Excel spreadsheet. To do this, Save As... > File Format > .csv

