---
title: "Variables_assignment_part2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# First, lets start with some quick review

Where are the environment, files, console, and script/markdown windows respectively?
How do these relate to our cooking analogy?

## R as a calculator
```{r}
#How would I get an output of the value two times 6 all to the power of three? Hint, PEMDAS
(2*6)^3


```
## Comparisons
```{r}
1 >= 2 #What does '>=' mean? Is this True or False?

15 <= 30 #What does '<=' mean? True or False?

15 == 2 #What does '==' mean? How is this different from just '='? True or False?

##Bonus! What data type is this output?
```
## Variable assignment 
```{r}
#What is the assignment operator in R? How would I go about assigning the value u the numeric value 2? 
u<-2


```

## Data types
```{r}
#What were the three basic data types? 
#character string
w <- "word"
interger <- 1
#What are some examples of these?
```


A quick sidenote on variable names: They can be anything! With a few exceptions:
do not use . in front of the variable, that'll store it as a hidden variable (not usually useful). They must also start with a letter of some sort
```{r}
_apple <- "fruit" #Will not work if prefixed with a number!

#Underscores are a great way to put spaces in long variable names

paired_t_test <- "statistical test"

```




## Vectorization: Vectors in R describes a set of values in a certain order of the same data type. This lets variables and functions have vectors for values. 
```{r}
1:8 #This is a vector of 1-8

#You can also create a more advanced vector 
2^(1:8)

#If you wanted to be more specific in your integer selection inside the vector, you can utilize the concatenate fxn. c() 
vector <- c(1,2,5,4)

vector
#You can also store characterstrings using this
vector_characters <- c("Red", "Blue", "Green")
vector_characters 

#Or both
vector_both <- c("Red", 1.2, "Purple")
vector_both

#Though it'll store these all as characters, which isn't useful for some applications. 
typeof(vector_both)
```



Challenge #2: Create a variable with a name of your own choosing that includes a vector of any 5 odd integers
```{r}
blah <- c(3,5,7,9,11)
```

