---
title: "Variables_assignment_part2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Vectorization: Vectors in R describes a set of values in a certain order of the same data type. This lets variables and functions have vectors for values. 
```{r}
1:8 #This is a vector of 1-8

#You can also create a more advanced vector 
2^(1:8)

#If you wanted to be more specific in your integer selection inside the vector, you can utilize the concatenate fxn. c() 
vector <- c(1,2,5,4)

vector
#You can also store characterstrings using this
vector_characters <- c("Red", "Blue", "Green")
vector_characters 

#Or both
vector_both <- c("Red", 1.2, "Purple")
vector_both

#Though it'll store these all as characters, which isn't useful for some applications. 
typeof(vector_both)
```

A quick sidenote on variable names: They can be anything! With a few exceptions:
do not use . in front of the variable, that'll store it as a hidden variable (not usually useful). They must also start with a letter of some sort
```{r}
apple <- "fruit" #Will not work if prefixed with a number!

#Underscores are a great way to put spaces in long variable names

paired_t_test <- "statistical test"

```

Challenge #2: Create a variable with a name of your own choosing that includes a vector of any 5 odd integers
```{r}

```

