---
title: "Data Types Intermediate Worksheet"
author: "Ruchira Krishnamurthy"
date: "1/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Different Data types: character strings, integers, and logical. 
## Character strings are composed of stringed together characteristics (so words)
## Integers are composed of only numbers
## Logical are TRUE/FALSE statements, this is important for understanding inputs for functions as well as their outputs. 
```{r}
#What is this data type?
a <- "apple"
# How can we check?
is.character(a)  
is.numeric(a) 
is.logical(a)
```


```{r}
#What is this data type?
num <- 4/10
is.character(num)
is.numeric(num) 
is.logical(num)
```

```{r}
# What is this data type?
f <- FALSE
is.character(f)
is.numeric(f)
is.logical(f) 
```

Vectorization: Vectors in R describes a set of values in a certain order of the same data type. This lets variables and functions have vectors for values. 
```{r}
# One way to create a vector is the concatenate function c()
vector <- c()
vector

# Example of a calculation you can do with vectors 

#You can also store character strings using the concatenate function
# Character vector with three movies
vector_characters <- c()
vector_characters 

# Or both
# 2 numbers + a movie
vector_both <- c()
vector_both
```
