---
title: "For Loops - Exercises 1 & 2"
author: "Joe Celebrezze"
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TEMPLATE

## Exercise 1
* Write a `for()` loop that (1) iterates over the numbers 1 to 10, (2) calculates the cube of each number and (3) saves the output in a storage vector. 
```{r}

```

## Exercise 2

* Write a `for()` loop that (1) iterates over the numbers 1 to 10, (2) adds each variable to the variable that succeeds it in vector x (e.g., 1 + 2; 2 + 3; etc.) and (3) saves the output in a storage vector.
```{r}

```











The key to the above exercises is below








# KEY

## Exercise 1

* Write a `for()` loop that (1) iterates over the numbers 1 to 10, (2) calculates the cube of each number and (3) saves the output in a storage vector. 
```{r}
x1 <-  c(1:10)
x1
y1 <- rep(NaN, length(x1))
y1

for (i in 1:length(x1)){
  y1[i] <- x1[i]^3
}

y1
```

## Exercise 2

* Write a `for()` loop that (1) iterates over the numbers 1 to 10, (2) adds each variable to the variable that succeeds it in vector x (e.g., 1 + 2; 2 + 3; etc.) and (3) saves the output in a storage vector.
```{r}
x2 <- c(1:10)
y2 <- rep(NaN, length(x2)-1)

for(i in 1:length(x2)-1){
  y2[i]<- x2[i]+x2[i+1]
}

y2
```