---
title: "intermediate review - publishable figures"
author: "Joe Celebrezze"
date: "1/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# For ggplot2, dplyr, etc.
library(tidyverse)

# For ggpubr section
library(ggpubr)

# For pretty statistical summary table section
library(gt)
library(gtsummary)

# For palmer penguins dataset
library(palmerpenguins) # This will load the palmer penguins dataset, as a dataset called penguins
```

```{r}
# view() to view the data set

```


The following script is taken and derived from Sam Sambado's "Publishable Figures" tutorial

# GGPlot2: An example of making nice looking plots

## First, let's set up a basic dotplot faceted by island and sex
- Numerical variables that we are interested in are flipper length and bill depth 
- Let's also color by species
```{r}
plot <- ggplot(data = , # omit NAs from dataset for visualization purposes
       aes()) + # choose our variables of interest
  geom_point() + # geom_point makes sense since there's 2 numeric values
  facet_grid() + # want to create a "grid" for our geom_points. facet_wrap is another option for 1 variable
  theme_bw() # this sets the entire theme to a white background with light bordering - play around with theme_*()
plot
```


## Let's fix some of the more clunky features from the above plot to make it look nicer
What are some features that we want to change?

```{r}
# We want to change lowercase "female" and "male" to uppercase
# State your new labels in a vector  
supp.labs <- c() 

# Add labels or names to the vector
names(supp.labs) <- c("female", "male")

supp.labs
```


```{r}
# Create an improved plot called "plot.improved"
# For colors: https://www.nceas.ucsb.edu/sites/default/files/2020-04/colorPaletteCheatsheet.pdf

```


## Now, messing with the theme a bit to get it looking exactly how we want it to.
1) Bold text for axes labels Bill and Flipper 
2) Change the colors of the facet grid rectangles 
3) Change the size of font in facet grid labels and make them bold   
4) Move the legend to the bottom of the graph  
5) Make legend text same size as facet_grid text
6) Make legend title a bold font
```{r}
# Create a more improved plot called plot.improved2
plot.improved2 <- plot.improved +
theme(axis.title = , # gives us bold text for axes labels Body & Flipper, 
        strip.background = , # changes the colors of the facet_grid rectangles
        strip.text = , 
        legend.position = , # move legend to bottom of graph
        legend.text = , # make legend text same size as facet_grid text
        legend.title = ) # make legend title a bold font

plot.improved2
```

