---
title: "intermediate review - publishable figures"
author: "Joe Celebrezze"
date: "1/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# For ggplot2, dplyr, etc.
library(tidyverse)

# For ggpubr section
library(ggpubr)

# For pretty statistical summary table section
library(gt)
library(gtsummary)

# For palmer penguins dataset
library(palmerpenguins)
```

The following script is taken and derived from Sam Sambado's "Publishable Figures" tutorial

# GGPlot2: An example of making nice looking plots

## First, let's set up a basic dotplot faceted by island and sex
```{r}
plot <- ggplot(data = na.omit(penguins), # omit NAs from dataset for visualization purposes
       aes(x = flipper_length_mm, y = bill_depth_mm, color = species)) + # choose our variables of interest
  geom_point() + # geom_point makes sense since there's 2 numeric values
  facet_grid(island ~ sex) + # want to create a "grid" for our geom_points. facet_wrap is another option for 1 variable
  theme_bw()

plot
```

## Let's fix some of the more clunky features from the above plot to make it look nicer
```{r}
supp.labs <- c("Female", "Male") # state your new labels 
names(supp.labs) <- c("female", "male")

plot.improved <- ggplot(data = na.omit(penguins), # omit NAs from dataset for visualization purposes
       aes(x = flipper_length_mm, y = bill_depth_mm, color = species)) + # choose our variables of interest
  geom_point() + # geom_point makes sense since there's 2 numeric values
  facet_grid(island ~ sex, # want to create a "grid" for our geom_points. facet_wrap is another option for 1 variable
             scale = "free_y",  # want to y axis to best fit each group, not for entire dataset
             labeller = labeller(sex = supp.labs)) + # want to change labels from lower to uppercase
  scale_color_manual(values = c("orange", "red", "blue")) + # choose colors manually (make sure use scale_COLOR not scale_FILL since in line 66 we said COLOR = species)
  labs(x = "Flipper Length (mm)", # change the x axis labels to something nicer to look at
       y = "Bill Depth (mm)", # change the y axis labels to something nicer to look at
       color = "Species") + # this changes the title of your legend
  theme_bw() # this sets the entire theme to a white background with light bordering - play around with theme_*()

plot.improved
```

## Now, messing with the theme a bit to get it looking exactly how we want it to.
```{r}
plot.improved2 <- plot.improved +
theme(axis.title = element_text(face = "bold"), # gives us bold text for axes labels Body & Flipper
        strip.background = element_rect(fill = "gray95"), # changes the colors of the facet_grid rectangles
        strip.text = element_text(size = 10, face = "bold" ), # change size of font in facet_grid label and make it bold
        legend.position = "bottom", # move legend to bottom of graph (there are many ways of doing this, this is one)
        legend.text = element_text(size = 10), # make legend text same size as facet_grid text
        legend.title = element_text(face = "bold")) # make legend title a bold font

plot.improved2
```

# Working with ggpubr package
With the ggpubr package, you can easily add things atop ggplot2 visualizations. Below is an example of adding p-values for comparisons between the penguin species, but you can also add things like:
  -  Adding background images
  -  Adding summary statistics
  -  Arranging multiple ggplots
  -  Visualizing error
  -  Other themes not included in ggplot2 package
  -  Etc.

```{r}
# state which comparisons you would like to make between the different levels of your factor `species`

my_comparisons <- list(c("Adelie", "Chinstrap"), c("Adelie", "Gentoo"), c("Chinstrap", "Gentoo"))

# slightly different format that ggplot() but relatively similar concepts
ggboxplot(na.omit(penguins), x = "species", y = "body_mass_g",  # data you want to plot
         add = "boxplot", add.params = list(fill = "grey89")) + # fill in boxplots
  stat_compare_means(comparisons = my_comparisons, aes(label = "p.signif")) + # print p-value of your comparisons
  stat_compare_means(label.y = 7500, label.x = .7) + # print statistical test used to make comparisons
  labs(x = "Species", # change axes labels
       y = "Body Mass (g)")
```

# Making a pretty statistical summary table
```{r}
penguins %>% 
  tbl_summary() #This looks okay, but it is a bit messy and includes data that is not overly statistically interesting. I think we can do better.

penguins_table <- penguins %>% 
  tbl_summary(by = species, # We want to group by species
              statistic = list(all_continuous() ~ "{mean} ({sd})"), # For all continuous variables, we will add the mean and the standard deviation in our specified format
              missing_text = "(Missing)", # To specify our missing data
  add_p() %>% # To add p value 
  add_stat_label(location = "column") # To specify on the table what stats we are collecting for each row
  
  #MESS WITH THE LABELS!
  
penguins_table
```

