---
title: "Variables_Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# You can store a variable using the assignment operator <-. These variables are then stored in your enviroment. 
```{r}
x <- 1

print(x) #Prints an out put of the variable inside the functions argument.

y = 2 #You can use the '=' as an assignment operator, however this is not recommended because many arguments are structured to use '=' inside a function, therefore its best practice in R to use '<-'

y #Just putting the variable will print the variable in your console 
```
# You can use variables inside other functions to make your code more 'dynamic' (meaning that you can change your variable earlier in your code and it'll subsequently change the rest of your outputs)
```{r}
z <- log(x) #Use the function log on variable x, this output will then be stored as a new variable called 'z' 

z + 15 #Add 15 to this new variable 

#Notice that if you didnt use variable assignments you would have to manually type in the outputs from your console to calculate what the log(x) + 15 is in this format. This 'dynamic' way of writing your code allows for clarity as well as a fluid method for changing your inputs. 

```
# You can also update your variables in your environment if you wanted to change them (log transformations are common here)
```{r}
x <- x + 2 #Notice that the variable 'x' updated from 2 to 3 in your environment
```

# Challenge Problem #1: Create a variable called u, make that variable be the value of x times 1.5
```{r}
u <- x * 1.5
```

# Data types: character strings, integers, and logical. 
## Character strings are composed of stringed together characteristics (so words)
## Integers are composed of only numbers
## Logical are TRUE/FALSE statements, this is important for understanding inputs for functions as well as their outputs. 
```{r}
#What is this data type?
b <- "banana" 
is.character(b)  
is.numeric(b) 
is.logical(b)
```
```{r}
#What is this data type?
nu <- 3/22
is.character(nu)
is.numeric(nu) 
is.logical(nu)
```

```{r}
#What is this data type?
t <- TRUE
is.character(t)
is.numeric(t)
is.logical(t) 
```

Vectorization: Vectors in R describes a set of values in a certain order of the same data type. This lets variables and functions have vectors for values. 
```{r}
1:8 #This is a vector of 1-8

#You can also create a more advanced vector 
2^(1:8)

#If you wanted to be more specific in your integer selection inside the vector, you can utilize the concatenate fxn. c() 
vector <- c(1,2,5,4)

vector
#You can also store characterstrings using this
vector_characters <- c("Red", "Blue", "Green")
vector_characters 

#Or both
vector_both <- c("Red", 1.2, "Purple")
vector_both

#Though it'll store these all as characters, which isn't useful for some applications. 
typeof(vector_both)
```

A quick sidenote on variable names: They can be anything! With a few exceptions:
do not use . in front of the variable, that'll store it as a hidden variable (not usually useful). They must also start with a letter of some sort
```{r}
apple <- "fruit" #Will not work if prefixed with a number!

#Underscores are a great way to put spaces in long variable names

paired_t_test <- "statistical test"

```

Challenge #2: Create a variable with a name of your own choosing that includes a vector of any 5 odd integers
```{r}

```

