---
title: "Allison Horst Looping Over Columns and Rows"
author: "Joe Celebrezze"
date: "2/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is all from:
https://allisonhorst.github.io/danc_loops/#

Made by the wonderful Allison Horst
https://www.allisonhorst.com/

# Necessary Packages
```{r}
library(palmerpenguins)
library(tidyverse)
View(Loblolly)
```

# ACROSS FUNCTION

## Introduction to "Across": Getting Means
```{r}
mean(penguins$bill_length_mm)

#What if we wanted to know TWO means GROUPED BY SPECIES?
penguins %>% 
  group_by(species) %>%
  summarize(mean.bill.length = mean(bill_length_mm, na.rm = TRUE),
            mean.bill.depth = mean(bill_depth_mm, na.rm = TRUE))
```
But, what if we wanted to know A LOT of means?
Use across to apply a function to many columns
  The default of across is to apply to ALL of the columns

## Example 1 - Across Function
```{r}
penguins.chr <- penguins %>% 
  mutate(across(.cols = everything(), # .cols = everything() is default
                as.character)) #function you want to apply

glimpse(penguins.chr) #taking a look at our new dataset
```

## Example 2 - Across Function, but Where?
```{r}
penguins %>% 
  mutate(across(.cols = where(is.factor), # Now, we only want to convert columns with factors
                tolower)) #function you want to apply
```

## Example 3 - Across Function, with Summarize, with helper functions
```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(across(ends_with("mm"), # We only want to apply to columns that end with 'mm'
                   mean, na.rm = TRUE)) # Function(s) we want to apply
```

# ROWWISE FUNCTION

## Example 5 (in handout) - Rowwise Function, Calculating a sum of all <dbl> columns by row
```{r}
penguins_ex5 <- penguins %>% 
  rowwise() %>%  # Applied by ROW
  mutate(sum_bill = sum(across(where(is.double)))) # Using Across()

glimpse(penguins_ex5)
```



