---
title: "Functions and Conditionals KEY"
author: "DAnC Team"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Necessary Packages
```{r}
#install.packages("tidyverse")
library(tidyverse)
```

# Introduction to Conditionals (JT)

# Conditionals in For Loops (Ruchira)

# Defining and Practicing Functions (Andrea)

Today we will be going over how to create your own functions in R!

Our first example here is a function to convert Fahrenheit to Celsius
```{r}
#Let's take a look at the parts of the function 
fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9
  return(temp_C)
}

#What is the temperature today?
fahrenheit_to_celsius()
```

Let's create a function to calculate the volume of a cylinder:

```{r}
#for everyone who has forgotten middle school math:
#pi*radius^2*height

cylinder_volume <- function(r,h) {
  volume <- pi*r^2*h
  return(volume)
}
#in this answer, we use r as the radius and h as the height

cylinder_volume(5,6)


```

Setting a default 

```{r}
#What if you wanted to set your cylinder volume function to have a default of 1?

cylinder_volume2 <- function(r, h=1) {
  volume <- pi*r^2*h
  return(volume)
}

cylinder_volume2(3)

#even though there was no input for the height the function works because a default of 1 is set 

```


Make your own zodiac sign generator!

```{r}
#for this example you will put in your birth year and get your zodiac animal

zodiac_sign <- function(year) {
  x <- year %% 12
  
  animals <- c("rat", "ox", "tiger", "rabbit", "dragon", "snake", "horse", "goat", "monkey", "rooster", "dog", "pig")
  
 return(animals[x-3])
}

zodiac_sign(1999)


```


# Bringing it all together (Joe)

Working off of the 'trees' dataset, what if we wanted to write a function to try and predict one of the variables (for example the height of the trees) using other variables?

Well, first we would have to decide *how* we want to predict the height of the trees. For the sake of ease, lets use a linear model and first try and predict the height of the trees using both the girth (diameter) and the volume of the trees.

Before we do anything else, it is always good practice to look at the dataset we are working with and do a quick exploratory visualization.

## Data Exploration
```{r}
trees

ggplot(trees, aes(x = Volume, y = Height, color = Girth)) + # Note: R is case-sensitive
  geom_point() + # For a scatterplot
  geom_smooth(se = FALSE, method = "lm", color = 'black') # Overlaying a linear model without the standard error bar (and specifying the color to black)
```

From the above data visualization, we see that the height is at least somewhat related to the girth and the volume of the tree, so using those to predict tree height is not completely out-of-the-question. If you were doing this for research purposes, you would do some more statistics; however, we are just doing this to learn about functions and conditionals.

But, we are not quite to the point of functions and conditionals yet... First, we need to make a linear model.

## Linear Model (and adding predicted height to our dataset)
```{r}
height.model <- lm(data = trees, Height ~ Girth + Volume) # This is one way to make a linear model where you specify the dataset, and then specify your explanatory variables (girth and volume) and your response variable (height)
summary(height.model) # This summarizes that linear model; WE WILL USE THIS LATER

height.predict <- cbind(trees, predict(height.model)) # To get predicted values of height from the linear model
trees$Predicted.Height <- height.predict$`predict(height.model)` # To add those predicted values to our dataset
```

Okay! Now, we are finally ready to make a function to predict the height using the girth and the volume

## Function with Girth and Volume Predicting Height
```{r}
predict.height <- function(girth, volume) {
  Predicted.Height <- 83.2958 - (1.8615*girth) + (0.5756*volume) # We get these values from the summary(height.model) above!
  return(Predicted.Height)
}

predict.height(girth = 8.3, volume = 10.3)

# To check out how our function did
trees
```

So, that function did pretty good, but I think we could add some cool things to it. What if we wanted the predicted height to be an output of inches, meters or cm instead of feet?

## Function Involving Conditionals -- Units!
```{r}
predict.height <- function(girth, volume, units = "feet") # units = "feet" in the parentheses of the funtion sets feet as the default (if you do not specify any units, it will return in feet)
  { 
  Predicted.Height <- 83.2958 - (1.8615*girth) + (0.5756*volume)
  if (units == "in"){
    return(Predicted.Height*12) # Alter the predicted height based on the conversion factors for the units
  }
  else if (units == "m"){
    return(Predicted.Height*0.3048)
  }
  else if (units == "cm") {
    return(Predicted.Height*30.48)
  }
  else{
    return(Predicted.Height)
  }
}

predict.height(8.3, 10.3, units = "m")

#Check our work!
predict.height(8.3, 10.3) * 0.3048 # See, here we do not specify units, but the output would be in feet because we specified feet as the default
```

One slight issue...
```{r}
predict.height(8.3, 10.3, units = "meters")
```
This will return the height in FEET because of how our function is structured. Think about it like this: if units are "in" then this, if units are "m" then this, if units are "cm" then this, in ANY OTHER CASE return this... Instead, we would probably want to use an else if statement for the feet and then use else to return an error so that if the user of the function (us) put in "meters" instead of "m" it would return an error letting us know that we did something wrong.

Nice, this function is becoming a bit more complex, but what if we wanted to predict height, but we could only measure either girth OR volume and we could not measure both?

Let's try it out
```{r}
# predict.height(girth = 8.3)
# our function does not like this... error pops up!
```

Clearly we need to alter our function. Conditionals could help with this too!

First, we need to make some new linear models that do not rely on either the girth or the volume (one model using only volume as an explanatory variable and one using only girth as an explanatory variable).
```{r}
height.model.no.girth <- lm(data = trees, Height ~ Volume)
summary(height.model.no.girth)

height.model.no.volume <- lm(data = trees, Height ~ Girth)
summary(height.model.no.volume)
```

## Function Using Conditionals Part 2 -- More Models!
```{r}
predict.height <- function(girth, volume, units = "feet") {
  if(girth == "no"){
    Predicted.Height <- 69.00336 + (0.2319*volume)
  }
  else if(volume == "no") {
    Predicted.Height <- 62.0313 + (1.0544*girth)
  }
  else{
    Predicted.Height <- 83.2958 - (1.8615*girth) + (0.5756*volume)
  }
  if (units == "inches"){
    return(Predicted.Height*12)
  }
  else if (units == "m"){
    return(Predicted.Height*0.3048)
  }
  else if (units == "cm") {
    return(Predicted.Height*30.48)
  }
  else{
    return(Predicted.Height)
  }
}

predict.height(girth = "no", 20)
predict.height(20, girth = "no") # if you define one of the variables (and units is the default in this case, then it does not matter which order you place the variables)
```

