---
title: "Functions and Conditionals TEMPLATE"
author: "DAnC Team"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Necessary Packages
```{r}
#install.packages("tidyverse")
library(tidyverse)
```

# Introduction to Conditionals (JT)

# Conditionals in For Loops (Ruchira)
Lets implement both for loops and conditionals in an example!
As a reminder:
* The basic structure of a for loop is: 

for (variable in a collection of variables) {  
expression(s)  
}  

Let's look at an example:
```{r}
# Let's create a dataset 
dataset <- data.frame(A = c(1:10), B= c(11:20))

# View dataset
dataset
```

Let's create a new column called C that will return "yes" or "no" depending on if the values of column B are greater than 15 (value "yes") or not (value "no"). 
```{r}
for(){
  if(){}
  else{}
}

dataset
```

Next let's look at a dataset from the "dataset" package called trees
```{r}                            
# Let's view a portion of our dataset first.
trees_data <- trees
head(trees_data)
```



Exercise:
Look through the height column of trees_data, and if the tree is larger than 73 inches, assign the value "tall" to a new column "Description", else assing the value "short" in the column "Description". View trees_data after to check that your code has worked. 

```{r}

```

# Defining and Practicing Functions (Andrea)
Today we will be going over how to create your own functions in R!

Our first example here is a function to convert Fahrenheit to Celsius
```{r}
#Let's take a look at the parts of the function 
fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9
  return(temp_C)
}

#What is the temperature today?
fahrenheit_to_celsius()
```

Let's create a function to calculate the volume of a cylinder:

```{r}
#for everyone who has forgotten middle school math:
#pi*radius^2*height

cylinder_volume <- function( , ) {
  volume <- 
  return(volume)
}
```
Setting a default 
```{r}
#What happens when we run the function with only one number?
cylinder_volume()

#How do we set a default?

```

Now let's create a new function from scratch!

Our "star sign" generator

```{r}

```




# Bringing it all together (Joe)

Working off of the 'trees' dataset, what if we wanted to write a function to try and predict one of the variables (for example the height of the trees) using other variables?

Well, first we would have to decide *how* we want to predict the height of the trees. For the sake of ease, lets use a linear model and first try and predict the height of the trees using both the girth (diameter) and the volume of the trees.

Before we do anything else, it is always good practice to look at the dataset we are working with and do a quick exploratory visualization.

## Data Exploration
```{r}
trees

ggplot(trees, aes(x = Volume, y = Height, color = Girth)) + # Note: R is case-sensitive
  geom_point() + # For a scatterplot
  geom_smooth(se = FALSE, method = "lm", color = 'black') # Overlaying a linear model without the standard error bar (and specifying the color to black)
```

From the above data visualization, we see that the height is at least somewhat related to the girth and the volume of the tree, so using those to predict tree height is not completely out-of-the-question. If you were doing this for research purposes, you would do some more statistics; however, we are just doing this to learn about functions and conditionals.

But, we are not quite to the point of functions and conditionals yet... First, we need to make a linear model.

## Linear Model (and adding predicted height to our dataset)
```{r}
height.model <- lm(data = trees, Height ~ Girth + Volume) # This is one way to make a linear model where you specify the dataset, and then specify your explanatory variables (girth and volume) and your response variable (height)
summary(height.model) # This summarizes that linear model; WE WILL USE THIS LATER

height.predict <- cbind(trees, predict(height.model)) # To get predicted values of height from the linear model
trees$Predicted.Height <- height.predict$`predict(height.model)` # To add those predicted values to our dataset
```

Okay! Now, we are finally ready to make a function to predict the height using the girth and the volume

## Function with Girth and Volume Predicting Height
```{r}

```

So, that function did pretty good, but I think we could add some cool things to it. What if we wanted the predicted height to be an output of inches, meters or cm instead of feet?

## Function Involving Conditionals -- Units!
```{r}

# Check our work!

```

One slight issue though...
```{r}

```


Nice, this function is becoming a bit more complex, but what if we wanted to predict height, but we could only measure either girth OR volume and we could not measure both?

Let's try it out
```{r}

```

Clearly we need to alter our function. Conditionals could help with this too!

First, we need to make some new linear models that do not rely on either the girth or the volume (one model using only volume as an explanatory variable and one using only girth as an explanatory variable).
```{r}
height.model.no.girth <- lm(data = trees, Height ~ Volume)
summary(height.model.no.girth)

height.model.no.volume <- lm(data = trees, Height ~ Girth)
summary(height.model.no.volume)
```

## Function Using Conditionals Part 2 -- More Models!
```{r}

# Check our work

```
