---
title: "Dplyr_mutate_filter"
author: "Andrea Chandler"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We'll be practicing with an R markdown file, remember you can make a new r chunk using the green button in the right hand corner or using the shortcut "ctrl alt I" for Windows and "Cmd Option I" for Mac. 


Downloading packages
```{r}
#First, download the necessary packages!
#install the palmerpenguin dataset, there are two penguin data sets, the "raw" data set and a simplified version we will be using today
install.packages("palmerpenguins")
#also install tidyverse if you have not done so already!
install.packages("tidyverse")

```


Calling packages 
```{r}
#use the library function to tell R which packages you want to use in this markdown file!

#the penguins data set is just called 'penguins' once you install it
library(palmerpenguins)

library(tidyverse)

```

First, let's briefly visualize and talk about our data!

```{r}
#use the View function to see your data in another tab
view(penguins)

#can also check out the other penguin data set!
penguins_raw
```

We will be learning the summarize function later, which can also help you get a sense of what your data looks like!


First, the filter function!

```{r}
#the general format of a function

#function(argument)

#Example of the filter function: filter by island - remember to specify your data set!

dplyr::filter(penguins, island == "Dream")

```

Pick a favorite penguin to filter by!
```{r}
#filter by chosen penguin
dplyr::filter(penguins, species == "Adelie")

```

Ok, but what if I have TWO favorite penguins
```{r}
#using the | operator
dplyr::filter(penguins, species == "Chinstrap" | species == "Gentoo")
#using the %in% operator
dplyr::filter(penguins, species %in% c("Chinstrap", "Gentoo"))
```


More filter practice - using <, >, >= and <= 
```{r}
penguins %>% 
  filter(flipper_length_mm > 200)

```

Now for our next function - mutate

```{r}
#mutate allows you to create a new column that is a modification of an old one

#for example, if you wanted a 'bill ratio' column, with bill ratio = bill length/bill depth
penguins %>% 
  mutate(bill_ratio = bill_length_mm/bill_depth_mm)





```

To note:
Renaming the column the same name will replace it!!