---
title: "Dplyr_mutate_filter"
author: "Andrea Chandler"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We'll be practicing with an R markdown file, remember you can make a new r chunk using the green button in the right hand corner or using the shortcut "ctrl alt I" for Windows and "Cmd Option I" for Mac. 


Downloading packages
```{r}
#First, download the necessary packages!
#install the palmerpenguin dataset, there are two penguin data sets, the "raw" data set and a simplified version we will be using today
#install.packages("palmerpenguins")
#also install tidyverse if you have not done so already!
#install.packages("tidyverse")

```


Calling packages 
```{r}
#use the library function to tell R which packages you want to use in this markdown file!

#the penguins data set is just called 'penguins' once you install it
library(palmerpenguins)

library(tidyverse)

```


```{r}
#type here for your code!
```



First, let's briefly visualize and talk about our data!

```{r}
#use the View function to see your data in another tab
View(penguins)


#can also check out the other penguin data set!

```

We will be learning the summarize function later, which can also help you get a sense of what your data looks like!


First, the filter function!

```{r}
#the general format of a function


#function(argument)

#Example of the filter function: filter by island Torgersen- remember to specify your data set!

filter(penguins, island == "Torgersen")


```

Pick a favorite penguin to filter by!
```{r}
#filter by chosen penguin

filter(penguins, species == "Adelie")

#FILTER BY CHINSTRAP

filter(penguins, species == "Chinstrap")


```

Ok, but what if I have TWO favorite penguins
```{r}
#using the | operator

filter(penguins, species == "Adelie" | species == "Chinstrap")

#using the %in% operator
filter(penguins, species %in% c("Adelie", "Chinstrap"))

```


More filter practice - using <, >, >= and <= 
```{r}

#filtering by penguins with body mass greater than or equal to 3000 (scientifically known as chonkers)

filter(penguins, body_mass_g >= 3000)

```

Now for our next function - mutate

```{r}
#mutate allows you to create a new column that is a modification of an old one

#mutate(dataset, new name = modifications of old columns)


#for example, if you wanted a 'bill ratio' column, with bill ratio = bill length/bill depth

mutate(penguins, bill_ratio = bill_length_mm/bill_depth_mm)


```

To note:
Renaming the column the same name will replace it!!
# Good note!