---
title: "dplr-Select-Group_by-Summarize Template"
author: "Ruchira Krishnamurthy"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# dplyr is a package in the tidyverse!
# install.packages("palmerpenguins")
# install.packages("tidyverse")
library(tidyverse)
library(palmerpenguins)
```

# Select
- select() allows us to pick or exclude certain variables in your data

# First let's look at the data set using head()
```{r view dataset}
# Assign data set "penguins" to the variable "penguins"
# View the first 6 rows

```

# Let's say we want to only view 3 columns from the penguin data and in a particular order.
```{r}


```

# View data set using head()
```{r}
# Note: the penguins data set does not change. Using %>% and select() we are choosing what columns or variables we would like to view. 


```

# View all the columns from species to bill_depth_mm as well as year

```{r}
# : operator

```

# Keep all the columns except the ones from species to bill_depth_mm and year
```{r}
# : operator + !or - operator

```


# Group_by + Summarize 
Use group_by() and summarize() to prepare a summary table of your data. 
- Group_by() takes an existing table and converts it into a grouped table where operations are performed "by group"
- Summarize() creates a new table containing a statistical summary of data you are interested in

# Example
Generate a summary table containing the mean of penguin body mass (body_mass_g), grouped by penguin species. Column in the summary table should be called mass_mean.
```{r}
# Note that the na.rm = TRUE argument is added to exclude missing values.

```

What other statistical operations can we do on this data?
```{r}

```

What other variables can we group by? 
```{r}

```

What happens if we didn't use the group_by() function?
```{r}

```


Useful links:
https://dplyr.tidyverse.org/reference/summarise.html#useful-functions
https://allisonhorst.shinyapps.io/dplyr-learnr/#section-welcome



