---
title: "ggplot2 quick introduction"
author: "Joe Celebrezze"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages, dataset
```{r}
#install.packages('tidyverse')
library(tidyverse)
library(palmerpenguins)
penguins <- penguins # This isn't really necessary, but it is nice to have the dataset up in your environment
```

# Looking at the dataset
```{r}
penguins
```

# Introduction to ggplot2

When plotting with ggplot2, you typically follow a certain syntax including geoms, stats, scales, themes, labels and facets (and other things not listed here, but conveniently listed on the ggplot2 cheatsheet here: https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf ). Let's go over *some* of these things without getting too caught up in the fine details.

Basic syntax to always start with: this will be your first line of code describing what dataset you're using and what variables you want to plot
  Example: ggplot(data = penguins, aes(x = bill_length_mm, y = body_mass_g)) where you specify your data and then the "aesthetics" of the graph (i.e. what is going to be on each axis)

Geoms: used to represent datapoints, variables
  Examples: geom_boxplot returns a boxplot, geom_point returns a scatterplot
  
Themes: used to change how your graph looks
  Examples: theme_bw(), theme_void(), theme_dark(), theme_linedraw(), theme_gray(), theme_minimal()
  
Facets: used to divide your plot into various subplots based on variables in your dataset
  Example: facet_wrap(~species) would return subplots for each species of penguin

# Let's make a plot
```{r}
ggplot(data = penguins, aes(x = species, y = flipper_length_mm))
```
Not a very exciting plot without datapoints represented. What feature should we add?

# Trial 2
```{r}
ggplot(data = penguins, aes(x = species, y = flipper_length_mm))+ #This '+' denotes that you're adding layers onto the graph above
  geom_boxplot()
```
Nice! We have a boxplot showing that Gentoo penguins have the largest flipper length on average. What else could we do?

- Separate by sex or site through a color

# Messing with the graph
```{r}
ggplot(data = penguins, aes(x = species, y = flipper_length_mm))+
  geom_boxplot()
```


# A little more in depth about what we can add to a ggplot
```{r}
penguins_noNA <- na.omit(penguins)

ggplot(penguins_noNA, aes(x = body_mass_g, y = bill_length_mm, color = species)) + #Notice that I did not include data = penguins_noNA; "data =" makes your code easier to understand, but is not absolutely necessary. The first argument in ggplot() is assumed to be the dataset even without "data ="
  geom_point()
```

```{r}
?geom_smooth

ggplot(penguins_noNA, aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, shape = "square", size = 2) +
  geom_smooth() #Default method is the Loess method. To change method use "method ="
```

```{r}
ggplot(penguins_noNA, aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, shape = "square", size = 1) +
  geom_smooth(method = 'lm', se = FALSE, size = 0.5) +
  facet_grid(island~sex) #In the format y ~ x where it facets on the y axis by island and on the x axis by sex

ggplot(penguins_noNA, aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, shape = "square", size = 1) +
  geom_smooth(method = 'lm', se = FALSE, size = 0.5) +
  facet_wrap(island~sex) #Facets by both, but facet_wrap just facets in a rectangular format and does not separate by axes like facet_grid
```

```{r}
ggplot(penguins_noNA, aes(x = body_mass_g, y = bill_length_mm, color = species)) +
  geom_point(alpha = 0.5, shape = "square", size = 1) +
  geom_smooth(method = 'lm', se = FALSE, size = 0.5) +
  facet_grid(island~sex) +
  theme_bw()
```

# Now, let's make our own graphs showing which penguins are the chunkiest

You can look at this by comparing species to body mass, flipper length to body mass, site to body mass. Any comparison to try and understand which penguins are the chunkiest.

```{r}

```

# But what if we want to customize how our graphs look

Axes titles, colors, etc.
JT will tell you how to do that!!
